---
title: "Data Preparation"
output: html_document

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'data_preparation.html'))})
---

#### Define Libraries
```{r echo=TRUE, message=FALSE, warning=FALSE, tidy=TRUE}
library("dplyr")
library("reshape2")
library("rmarkdown")
library("magrittr")
library("AER")
library("tidyverse")
```

#### Define Path
```{r tidy=TRUE}
dir.wrk <- getwd()
dir.data <- file.path(dir.wrk, "dataset/data")
dir.outdata <- file.path(dir.wrk, "dataset/outputdata")
dir.plots <- file.path(dir.wrk, "dataset/plots")
```

#### Define Files
```{r tidy=TRUE}
file.region <- file.path(dir.data, "classification_region.csv")
file.incomeclass <- file.path(dir.data, "classification_income.csv")
file.pop <- file.path(dir.data, "world_population_2022.csv")
file.gdp <- file.path(dir.data, "PC_GDP_2022.csv")
file.cons <- file.path(dir.data, "consumption_2022.csv")
```

#### Per capita GDP at 2017 constant dollar PPP (source: WDI updated March 28, 2024)
#### https://databank.worldbank.org/source/world-development-indicators#
```{r tidy=TRUE}
data_gdp <- read.csv(file.gdp, header=T, stringsAsFactors = FALSE)
data_gdp <- data_gdp %>% dplyr::select(1, 2, X2000:X2022)

data_gdp <- data_gdp %>% dplyr::rename(Country = Country.Name, Code = Country.Code)
data_gdp <- data_gdp %>% gather(Year, gdp, -Country, -Code)

data_gdp$Year <- stringr::str_replace(data_gdp$Year, "X", "")
data_gdp$Year <-as.numeric(data_gdp$Year)
```

#### Population data (Source:WDI)
```{r tidy=TRUE}
wpop <- read.csv(file.pop, header=T, stringsAsFactors = FALSE)
wpop <- wpop %>% select(1,2, X2000:X2022)
wpop <- na.omit(wpop)
wpop <- wpop %>% dplyr::rename(Country = Country.Name, Code = Country.Code)
wpop <- wpop %>% gather(Year, pop, -c(Country, Code))
wpop$Year <- stringr::str_replace(wpop$Year, "X", "")
wpop$Year <- as.numeric(wpop$Year)

gdp_pop <- inner_join(data_gdp, wpop, by = c("Country", "Code", "Year"))
gdp_pop$gdp <- as.numeric(gdp_pop$gdp)
gdp_pop <- na.omit(gdp_pop)
gdp_pop <- gdp_pop %>% filter(gdp != 0)
gdp_pop$pc_gdp <- gdp_pop$gdp/gdp_pop$pop
```

#### Country classification: Income (source: World Bank) 
#### https://datahelpdesk.worldbank.org/knowledgebase/articles/906519-world-bank-country-and-lending-groups)

```{r tidy=TRUE}
dat_class1 <- read.csv(file.incomeclass, header=T, stringsAsFactors = FALSE)
data_class1 <- dat_class1 %>% dplyr::select(1, 2, X2000:X2022)
```

#### Classification: region
```{r tidy=TRUE}
dat_class2 <- read.csv(file.region, header=T, stringsAsFactors = FALSE)
dat_class2 <- dat_class2 %>% dplyr::select(1:3)
```

#### join country classification data frames
```{r tidy=TRUE}
data_class <- left_join(dat_class2, data_class1, by = c("Code", "Country"))
dat_class <- data_class %>% gather(Year, in_class, -c(Code, Country, Region))
dat_class$Year <- stringr::str_replace(dat_class$Year, "X", "")
dat_class$Year <- as.numeric(dat_class$Year)
dat_class <- dat_class %>% select(1, 3, 4, 5)
```

#### Join Per Capita GDP, Population and Classification data frames
```{r tidy=TRUE}
gdp_pop <- list(dat_class, gdp_pop) %>% reduce(inner_join, by = c("Code", "Year"))

file.output <- file.path(dir.outdata, "gdp_pop.csv")
write.table(gdp_pop, file.output, sep=",", row.names=FALSE, col.names=TRUE, quote=FALSE)
```

#### Consumption Expenditure calculation from 2000 to 2022 (source: WDI, updated Feb 21, 2024)
```{r include=TRUE}
con_expd <- read.csv(file.cons, header=T, stringsAsFactors = FALSE)
con_expd <- con_expd %>% select(1, 2, X2000:X2022)
con_expd <- con_expd %>% dplyr::rename(Country = Country.Name, Code = Country.Code)
con_expd <- con_expd %>% gather(Year, expnd, -c(Code, Country))
con_expd$Year <- stringr::str_replace(con_expd$Year, "X", "")
con_expd$Year <- as.numeric(con_expd$Year)
```

#### Inner join consumption and population
```{r tidy=TRUE}
expd_pop <- inner_join(con_expd, wpop)
expd_pop$expnd <- as.numeric(expd_pop$expnd)
expd_pop <- na.omit(expd_pop)
con_pop <- expd_pop %>% filter(expnd != 0)

con_pop <- con_pop %>% mutate(pcon = expnd/pop)
con_pop <- con_pop %>% select(1:3,5,6)
```

#### Join Per Capita Expenditure, Population and Classification data frames
#### Join GDP, Consumption Expenditure, Population and Classification data frames
```{r tidy=TRUE}
con_pop <- list(dat_class, con_pop) %>% reduce(inner_join, by = c("Code", "Year"))
file.output <- file.path(dir.outdata, "con_pop.csv")
write.table(con_pop, file.output, sep=",", row.names=FALSE, col.names=TRUE, quote=FALSE)

dat_all <- list(con_pop, gdp_pop) %>% reduce(inner_join, by = c("Code", "Country", "Year", "pop", "Region", "in_class"))

file.output <- file.path(dir.outdata, "dat_all.csv")
write.table(dat_all, file.output, sep=",", row.names=FALSE, col.names=TRUE, quote=FALSE)
```