---
title: "Lorenz Kakwani"
output: html_document

knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'Lorenz Kakwani.html'))})
---


#### Define Libraries
```{r echo=TRUE, message=FALSE, warning=FALSE, tidy=TRUE}
library("dplyr")
library("reshape2")
library("rmarkdown")
library("magrittr")
library("AER")
library("tidyverse")
library("dineq")
library("reldist")
library("ineq")
library("ggpubr") # arrange graphs
library("IC2")
library("ggrepel")
```

#### Define Path
```{r tidy=TRUE}
dir.wrk <- getwd()
dir.data <- file.path(dir.wrk, "dataset/outputdata")
dir.outdata <- file.path(dir.wrk, "dataset/outputdata")
dir.plots <- file.path(dir.wrk, "dataset/plots")
```

#### Define Files
```{r tidy=TRUE}
file.dat_all <- file.path(dir.outdata, "dat_all.csv")
file.gdp_pop <- file.path(dir.outdata, "gdp_pop.csv")
```

#### main data
```{r tidy=TRUE}
gdp_con144 <- read.csv(file.dat_all, header=T, stringsAsFactors = FALSE)
gdp_con144 <- gdp_con144 %>% dplyr::select(1:7, 9)
```

#### GDP and Consumption (common 2019, 2020 and 2021)
```{r tidy=TRUE}
gdp_con19 <- gdp_con144 %>% filter(Year == 2019)
gdp_con20 <- gdp_con144 %>% filter(Year == 2020)
gdp_con21 <-gdp_con144 %>% filter(Year == 2021)

gdp_con19_code <- gdp_con19 %>% dplyr::select(1, 2)
gdp_con20_code <- gdp_con20 %>% dplyr::select(1, 2)
gdp_con21_code <- gdp_con21 %>% dplyr::select(1, 2)
code44 <- list(gdp_con19_code, gdp_con20_code, gdp_con21_code) %>% reduce(inner_join)
gdp_con44 <- inner_join(code44, gdp_con144)
gdp_con44$pop <-as.numeric(gdp_con44$pop)
```

#### Unweighted Gini Indices for Per Capita GDP and Consumption (for 144 Countries)
### preliminary set up for Lorenz and concentration curves (2019)

```{r tidy=TRUE}
## Ranked by 2019 GDP (So, GDP Lorenz and Consumption Concentration curve)
ge_2019 <- gdp_con44 %>% filter(Year==2019)

gg_19 <- ge_2019 %>% arrange(pc_gdp)  ## ranked by 2019 GDP
gg_19$one <- "1" 
gg_19$one <-as.numeric(gg_19$one)
gg_19$cum_unit = cumsum(gg_19$one)/sum(gg_19$one)  ## cum_unit ranked by 2019 per capita GDP

gg_19$cgdp = cumsum(gg_19$pc_gdp)/sum(gg_19$pc_gdp)      ## cumulative per capita GDP
gg_19$cpcon = cumsum(gg_19$pcon)/sum(gg_19$pcon)  ## cumulative per capita consumption
```

#### 2019 cumulative consumption ranked by 2019 consumption (Consumptiion lorenz)
```{r tidy=TRUE}
ge_19 <- ge_2019 %>% arrange(pcon)  ## Ranked by 2019 per capita consumption
ge_19$one <- "1" 
ge_19$one <- as.numeric(ge_19$one)
ge_19$cum_unt = cumsum(ge_19$one)/sum(ge_19$one)
ge_19$cpcon = cumsum(ge_19$pcon)/sum(ge_19$pcon)
```

#### 2020 cumulative data ranked by GDP
```{r tidy=TRUE}
ge_2020 <- gdp_con44 %>% filter(Year==2020)

gg_20 <- ge_2020 %>% arrange(pc_gdp)  ## ranked by 2020 GDP
gg_20$one <- "1" 
gg_20$one <-as.numeric(gg_20$one)
gg_20$cum_unit = cumsum(gg_20$one)/sum(gg_20$one)

gg_20$cgdp = cumsum(gg_20$pc_gdp)/sum(gg_20$pc_gdp)
gg_20$cpcon <-cumsum(gg_20$pcon)/sum(gg_20$pcon)
```

#### 2020 cumulative consmption ranked by 2020 consumption (Consumption Lorenz)
```{r tidy=TRUE}
ge_20 <- ge_2020 %>% arrange(pcon)  ## Ranked by 2020 consumption
ge_20$one <- "1" 
ge_20$one <- as.numeric(ge_20$one)
ge_20$cum_unt = cumsum(ge_20$one)/sum(ge_20$one)
ge_20$cpcon = cumsum(ge_20$pcon)/sum(ge_20$pcon)
```

#### 2021 cumulative data: ranked by 2021 GDP
```{r tidy=TRUE}
ge_2021 <- gdp_con44 %>% filter(Year==2021)

gg_21 <- ge_2021 %>% arrange(pc_gdp)  ## ranked by 2021 GDP
gg_21$one <- "1" 
gg_21$one <-as.numeric(gg_21$one)
gg_21$cum_unit = cumsum(gg_21$one)/sum(gg_21$one)

gg_21$cgdp = cumsum(gg_21$pc_gdp)/sum(gg_21$pc_gdp)
gg_21$cpcon <-cumsum(gg_21$pcon)/sum(gg_21$pcon)
```

#### 2021 cumulative consumption ranked by 2021 consumption (Consumption Lorenz) 
```{r tidy=TRUE}
ge_21 <- ge_2021 %>% arrange(pcon)  ## Ranked by 2021 consumption
ge_21$one <- "1" 
ge_21$one <- as.numeric(ge_21$one)
ge_21$cum_unt = cumsum(ge_21$one)/sum(ge_21$one)
ge_21$cpcon = cumsum(ge_21$pcon)/sum(ge_21$pcon)
```

#### Unweighted Concentration Curves for 2019 (library(ggrepel))
```{r tidy=TRUE}
gg_19$ccode <- gg_19$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
ge_19$ccode <- ge_19$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
gg_20$ccode <- gg_20$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
ge_20$ccode <- ge_20$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
gg_21$ccode <- gg_21$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
ge_21$ccode <- ge_21$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
```

#### Unweighted Weighted Gini
```{r tidy=TRUE}
gini144 <- gdp_con44 %>% group_by(Year) %>%
  summarize(gngdp = gini(pc_gdp), wgn_gdp = gini.wtd(pc_gdp, pop), gncon=gini(pcon), 
            wgn_con=gini.wtd(pcon, pop), n())
```

#### LAC values (Damgaard and Weiner: Describing Inequality in Plant Size or Fecundity, 2000. Ecology 81(4),1139-1142.library("ineq"))

```{r tidy=TRUE}
## Population weighted and unweighted LAC
LAC_gdpcon44 <- gdp_con44 %>% group_by(Year) %>%
  summarize(WLAC_gdp = round(Lasym(pc_gdp, pop), 3), WLAC_con = round(Lasym(pcon, pop), 3),
            ULAC_gdp = round(Lasym(pc_gdp), 3), ULAC_con = round(Lasym(pcon), 3)) %>% as.data.frame()
```

#### Unweighted Lorenz plots for 2019 to 2021
##### 2019
```{r tidy=TRUE}
gge19 <- ggplot() + 
  geom_abline(color = "grey")+
  geom_line(aes(x=ge_19$cum_unt, y=ge_19$cpcon, color = "Consumption (0.437, 0.866)"))+
  geom_line(aes(x=gg_19$cum_unit, y=gg_19$cgdp, color = "GDP (0.509, 0.874)"))+
  labs(x = "", y = "")+
  geom_text_repel(data = subset(ge_19, ccode),
                 mapping = aes(x= cum_unt, y=cpcon, label = Code), color = "green3", size = 2)+
  geom_text_repel(data = subset(gg_19, ccode),
                 mapping = aes(x =cum_unit, y=cgdp, label = Code), color = "red", size = 2) +
  scale_color_hue(name = "Unweighted (2019)", direction = -1)+
  theme(text=element_text(size=8))+
  theme(axis.title.x = element_text(size=10))+
  theme(axis.title.y = element_text(size=10))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(legend.position= c(0.40, 0.84))+
  theme(axis.title.x =element_blank())+
  theme(axis.title.y=element_blank())+
  theme(legend.title = element_text(size = 7),legend.text = element_text(size = 7))+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))
```

#### Unweighted Lorenz for 2020
```{r tidy=TRUE}
gge20 <- ggplot() + 
  geom_abline(color = "grey")+
  geom_line(aes(x=ge_20$cum_unt, y=ge_20$cpcon, color = "Consumption (0.429, 0.870)"))+
  geom_line(aes(x=gg_20$cum_unit, y=gg_20$cgdp, color = "GDP (0.504, 0.885)"))+
  labs(x = "",y = "")+
  geom_text_repel(data = subset(ge_20, ccode),
                  mapping = aes(x= cum_unt, y=cpcon, label = Code), color = "green3", size = 2) +
  geom_text_repel(data = subset(gg_20, ccode),
                  mapping = aes(x =cum_unit, y=cgdp, label = Code), color = "red", size = 2) +
  scale_color_hue(name = "Unweighted (2020)", direction = -1)+
  theme(text=element_text(size=8))+
  theme(axis.title.x = element_text(size=10))+
  theme(axis.text.y=element_blank())+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(legend.title = element_text(size = 7),legend.text = element_text(size = 7))+
  theme(legend.position= c(0.39, 0.84))+
  theme(axis.title.x =element_blank(), axis.ticks.y=element_blank())+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))
```

#### Unweighted Lorenz 2021
```{r tidy=TRUE}
gge21 <- ggplot() + 
  geom_abline(color = "grey")+
  geom_line(aes(x=ge_21$cum_unt, y=ge_21$cpcon, color = "Consumption (0.431, 0.847)"))+
  geom_line(aes(x=gg_21$cum_unit, y=gg_21$cgdp, color = "GDP (0.508, 0.881)"))+
  labs(x = "", y = "")+
  geom_text_repel(data = subset(ge_21, ccode),
                  mapping = aes(x= cum_unt, y=cpcon, label = Code), color = "green3", size = 2) +
  geom_text_repel(data = subset(gg_21, ccode),
                  mapping = aes(x =cum_unit, y=cgdp, label = Code), color = "red", size = 2) +
  scale_color_hue(name = "Unweighted (2021)", direction = -1)+
  theme(text=element_text(size=8))+
  theme(axis.title.x = element_text(size=10))+
  theme(axis.text.y=element_blank())+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(legend.title = element_text(size = 7),legend.text = element_text(size = 7))+
  theme(legend.position= c(0.39, 0.84))+
  theme(axis.title.x =element_blank(), axis.ticks.y=element_blank())+
  theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))
```

#### Unweighted Lorenz: 2019, 2020, and 2021
```{r tidy=TRUE}
ugg19_21 <-ggarrange(gge19, gge20, gge21, labels = c("a", "b", "c"), hjust = - 1, 
          font.label = list(size = 8, color = "black", face = "bold"), 
          ncol = 3, nrow = 1)
```

#### Preliminary set up for Weighted Lorenz and Concentration Curves
#### 2019 curves ranked by 2019 GDP
```{r tidy=TRUE}
wg_19 <- ge_2019 %>% arrange(pc_gdp)
wg_19$one <- "1" 
wg_19$one <-as.numeric(wg_19$one)
wg_19$cum_unit = cumsum(wg_19$one*wg_19$pop)/sum(wg_19$one*wg_19$pop)  ## cum_unit ranked by per capita GDP

wg_19$cgdp = cumsum(wg_19$pc_gdp*wg_19$pop)/sum(wg_19$pc_gdp*wg_19$pop)      ## cumulative per capita GDP
wg_19$cpcon = cumsum(wg_19$pcon*wg_19$pop)/sum(wg_19$pcon*wg_19$pop)
```

#### Cumulative 2019 Consumption, Ranked by 2019 Consumption
```{r tidy=TRUE}
we_19 <- ge_2019 %>% arrange(pcon)
we_19$one <- "1" 
we_19$one <- as.numeric(we_19$one)
we_19$cum_unt = cumsum(we_19$one*we_19$pop)/sum(we_19$one*we_19$pop) ## ranked by per capita consumption
we_19$cpcon = cumsum(we_19$pcon*we_19$pop)/sum(we_19$pcon*we_19$pop)  ## cumulative per capita consumption
```

#### Priliminary set up (2020) ranked by 2020 GDP
```{r tidy=TRUE}
wg_20 <- ge_2020 %>% arrange(pc_gdp)
wg_20$one <- "1" 
wg_20$one <-as.numeric(wg_20$one)
wg_20$cum_unit = cumsum(wg_20$one*wg_20$pop)/sum(wg_20$one*wg_20$pop)

wg_20$cgdp = cumsum(wg_20$pc_gdp*wg_20$pop)/sum(wg_20$pc_gdp*wg_20$pop)
wg_20$cpcon = cumsum(wg_20$pcon*wg_20$pop)/sum(wg_20$pcon*wg_20$pop)
```

#### Cumulative 2020 Consumption Ranked by 2020 Consumption 
```{r tidy=TRUE}
we_20 <- ge_2020 %>% arrange(pcon)
we_20$one <- "1" 
we_20$one <- as.numeric(we_20$one)
we_20$cum_unt = cumsum(we_20$one*we_20$pop)/sum(we_20$one*we_20$pop)
we_20$cpcon = cumsum(we_20$pcon*we_20$pop)/sum(we_20$pcon*we_20$pop)
```

#### Weighted Cumulative Values
#### Priliminary set up (2021) ranked by 2021 GDP
```{r tidy=TRUE}
wg_21 <- ge_2021 %>% arrange(pc_gdp)
wg_21$one <- "1" 
wg_21$one <-as.numeric(wg_21$one)
wg_21$cum_unit = cumsum(wg_21$one*wg_21$pop)/sum(wg_21$one*wg_21$pop)

wg_21$cgdp = cumsum(wg_21$pc_gdp*wg_21$pop)/sum(wg_21$pc_gdp*wg_21$pop)
wg_21$cpcon = cumsum(wg_21$pcon*wg_21$pop)/sum(wg_21$pcon*wg_21$pop) 
```

#### cumulative 2021 consumption ranked by 2021 consumption 
```{r tidy=TRUE}
we_21 <- ge_2021 %>% arrange(pcon)
we_21$one <- "1" 
we_21$one <- as.numeric(we_21$one)
we_21$cum_unt = cumsum(we_21$one*we_21$pop)/sum(we_21$one*we_21$pop)
we_21$cpcon = cumsum(we_21$pcon*we_21$pop)/sum(we_21$pcon*we_21$pop)
```

#### Lorenz annotations with a few major/populous countries
```{r tidy=TRUE}
wg_19$ccode <- wg_19$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
wg_20$ccode <- wg_20$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
wg_21$ccode <- wg_21$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
we_19$ccode <- we_19$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
we_20$ccode <- we_20$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
we_21$ccode <- we_21$Code %in% c("NGA", "BRA", "IND", "CHN", 
                                 "RUS", "GBR", "CAN", "USA")
```

#### Lorenz plots for consumption and income, countries ranked by respective year consumption or income 
### Plot for Population Weighted Lorenz (2019) 
```{r tidy=TRUE}
wgg19 <- ggplot() + 
  geom_abline(color = "grey")+
  geom_line(aes(x=we_19$cum_unt, y=we_19$cpcon, color = "Consumption (0.445, 1.180)"), size =0.8)+
  geom_line(aes(x=wg_19$cum_unit, y=wg_19$cgdp, color = "GDP (0.454, 1.174)"), size=0.8)+
  labs(x = "", y = "")+
  scale_color_hue(name = "Weighted (2019)", direction = -1)+
  geom_text_repel(data = subset(we_19, ccode),
                  mapping = aes(x=cum_unt,y=cpcon, label = Code), color = "green3", size = 2) +
  geom_text_repel(data = subset(wg_19, ccode),
                  mapping = aes(x=cum_unit,y=cgdp, label = Code), color = "red", size = 2) +
  theme(text=element_text(size=8))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(legend.title = element_text(size = 7),legend.text = element_text(size = 7))+
  theme(axis.title.x = element_text(size=10))+
  theme(axis.title.y = element_text(size=10)) +
  theme(axis.title.x=element_blank())+
  theme(axis.title.y=element_blank())+
  theme(legend.position= c(0.40, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))
```

#### Plot for Population Weighted Lorenz (2020)
```{r tidy=TRUE}
wgg20 <- ggplot() + 
  geom_abline(color = "grey")+
  geom_line(aes(x=we_20$cum_unt,y=we_20$cpcon, color = "Consumption (0.443, 1.155)"), size =0.8)+
  geom_line(aes(x=wg_20$cum_unit, y=wg_20$cgdp, color = "GDP (0.453, 1.187)"), size=0.8)+
  labs(x = "", y = "")+
  scale_color_hue(name = "Weighted (2020)", direction = -1)+
  geom_text_repel(data = subset(we_20, ccode),
                  mapping = aes(x=cum_unt, y=cpcon, label = Code), color = "green3", size = 2) +
  geom_text_repel(data = subset(wg_20, ccode),
                  mapping = aes(x =cum_unit, y=cgdp, label = Code), color = "red", size = 2) +
  theme(text=element_text(size=8))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(legend.title = element_text(size = 7),legend.text = element_text(size = 7))+
  theme(axis.text.y=element_blank(), axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank())+
  theme(legend.position= c(0.39, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))
```

#### Plot for Population Weighted Lorenz (2021)
```{r tidy=TRUE}
wgg21 <- ggplot() + 
  geom_abline(color = "grey")+
  geom_line(aes(x=we_21$cum_unt,y=we_21$cpcon, color = "Consumption (0.442, 1.147)"), size =0.8)+
  geom_line(aes(x=wg_21$cum_unit, y=wg_21$cgdp, color = "GDP (0.455, 1.208)"), size=0.8)+
  labs(x = "", y = "")+
  scale_color_hue(name = "Weighted (2021)", direction = -1)+
  geom_text_repel(data = subset(we_21, ccode),
                  mapping = aes(x=cum_unt, y=cpcon, label = Code), color = "green3", size = 2) +
  geom_text_repel(data = subset(wg_21, ccode),
                  mapping = aes(x =cum_unit, y=cgdp, label = Code), color = "red", size = 2) +
  theme(text=element_text(size=8))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(legend.title = element_text(size = 7),legend.text = element_text(size = 7))+
  theme(axis.title.y= element_blank(), axis.ticks.y=element_blank())+
  theme(axis.title.x=element_blank())+
  theme(legend.position= c(0.39, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))
```

#### plots 
```{r fig.height = 4.5, fig.width = 7, fig.align = "center"}
wgg19_21 <-ggarrange(wgg19, wgg20, wgg21, labels = c("d", "e", "f"), hjust= -1, vjust=0.5,
                     font.label = list(size = 8, color = "black", face = "bold"), 
                     ncol = 3, nrow = 1)

lorenz19_21 <-ggarrange(ugg19_21, wgg19_21, ncol = 1, nrow =2)

lorenz_1921 <- annotate_figure(lorenz19_21, left = text_grob("Cumulative % of GDP or Consumption Per Capita", rot = 90, vjust = 1, size =8), bottom = text_grob("Countries Ranked by GDP or Consumption Per Capita", vjust=0, size = 8))

file.plot <- file.path(dir.plots, "lorenz_1921.png")
png(file.plot, height=4.5, width=7, units="in", bg="white", res=300)
lorenz_1921
dev.off()
```
#### Concentration Values: ranked by corresponding year GDP
#### So for Per Capita consumption, the values are Pseudo Gini (IC2 Package)

```{r tidy=TRUE}
## unweighted p-seudo Gini
ci_con19 <- calcSConc(ge_2019[, "pcon"], ge_2019[,"pc_gdp"], param=2)$ineq$index
ci_con20 <- calcSConc(ge_2020[, "pcon"], ge_2020[,"pc_gdp"], param=2)$ineq$index
ci_con21 <- calcSConc(ge_2021[, "pcon"], ge_2021[,"pc_gdp"], param=2)$ineq$index

## Weighted p-seudo Gini
wci_con19 <- calcSConc(ge_2019[, "pcon"], ge_2019[,"pc_gdp"], w = ge_2019[,"pop"], param=2)$ineq$index
wci_con20 <- calcSConc(ge_2020[, "pcon"], ge_2020[,"pc_gdp"], w = ge_2020[,"pop"], param=2)$ineq$index
wci_con21 <- calcSConc(ge_2021[, "pcon"], ge_2021[,"pc_gdp"], w = ge_2021[,"pop"], param=2)$ineq$index
```

#### Concentration/Lorenz curves for consumption and income ranked by respective year's GDP
#### For 2019
```{r tidy=TRUE}
ucc19 <- ggplot(gg_19, aes(x=cum_unit)) + 
  geom_abline(color = "grey")+
  geom_line(aes(y=cgdp, color = "GDP (0.509)"))+
  geom_line(aes(y=cpcon, color = "Consumption (0.425)"))+
  labs(x = "", y = "")+
  geom_text_repel(data = subset(gg_19, ccode),
                  mapping = aes(y=cpcon, label = Code), color = "green3", size = 2) +
  geom_text_repel(data = subset(gg_19, ccode),
                  mapping = aes(y=cgdp, label = Code), color = "red", size = 2) +
  scale_color_hue(name = "Unweighted (2019)", direction = -1)+
  theme(legend.title = element_text(size = 8),legend.text = element_text(size = 7))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  theme(axis.title.x =element_blank())+
  theme(text=element_text(size=8))+
  theme(legend.position= c(0.38, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))
```

#### For 2020
```{r tidy=TRUE}
ucc20 <- ggplot(gg_20, aes(x=cum_unit)) + 
  geom_abline(color = "grey")+
  geom_line(aes(y=cpcon, color = "Consumption (0.418)"))+
  geom_line(aes(y=cgdp, color = "GDP (0.504)"))+
  labs(x = "", y = "")+
  geom_text_repel(data = subset(gg_20, ccode),
                  mapping = aes(y=cgdp, label = Code), color = "red", size = 2) +
  geom_text_repel(data = subset(gg_20, ccode),
                  mapping = aes(y=cpcon, label = Code), color = "green3", size = 2) +
  scale_color_hue(name= "Unweighted (2020)", direction = -1)+
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  theme(legend.title = element_text(size = 8),legend.text = element_text(size = 7))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(axis.title.x =element_blank())+
  theme(text=element_text(size=8))+
  theme(legend.position= c(0.36, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))
```

#### For 2021
```{r tidy=TRUE}
ucc21 <- ggplot(gg_21, aes(x=cum_unit)) + 
  geom_abline(color = "grey")+
  geom_line(aes(y=cpcon, color = "Consumption (0.420)"))+
  geom_line(aes(y=cgdp, color = "GDP (0.508)"))+
  labs(x = "", y = "")+
  geom_text_repel(data = subset(gg_21, ccode),
                  mapping = aes(y=cgdp, label = Code), color = "red", size = 2) +
  geom_text_repel(data = subset(gg_21, ccode),
                  mapping = aes(y=cpcon, label = Code), color = "green3", size = 2) +
  scale_color_hue(name= "Unweighted (2021)", direction = -1)+
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(axis.text.x = element_blank(), axis.ticks = element_blank()) +
  theme(legend.title = element_text(size = 8),legend.text = element_text(size = 7))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(axis.title.x =element_blank())+
  theme(text=element_text(size=8))+
  theme(legend.position= c(0.36, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))

ucc19_21 <- ggarrange(ucc19, ucc20, ucc21, labels = c("a", "b", "c"), ncol = 3, nrow = 1,
          hjust = -1.5, font.label = list(size = 8, color = "black", face = "bold"))
```

#### Weighted Concentration Plots ranked by corresponding year's per capita GDP
```{r tidy=TRUE}
wcc19 <- ggplot(wg_19, aes(x=cum_unit)) + 
  geom_abline(color = "grey")+
  geom_line(aes(y=cpcon, color = "Consumption (0.434)"))+
  geom_line(aes(y=cgdp, color = "GDP (0.454)"))+
  labs(x = "", y = "")+
  geom_text_repel(data = subset(wg_19, ccode),
                  mapping = aes(x=cum_unit, y=cgdp, label = Code), color = "red", size = 2) +
  geom_text_repel(data = subset(wg_19, ccode),
                  mapping = aes(x=cum_unit, y=cpcon, label = Code), color = "green3", size = 2) +
  scale_color_hue(name= "Weighted (2019)", direction = -1)+
  theme(axis.ticks.x = element_blank(), axis.ticks.y = element_blank()) +
  theme(legend.title = element_text(size = 8),legend.text = element_text(size = 7))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(axis.title.x =element_blank())+
  theme(text=element_text(size=8))+
  theme(legend.position= c(0.36, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))

wcc20 <- ggplot(wg_20, aes(x=cum_unit)) + 
  geom_abline(color = "grey")+
  geom_line(aes(y=cpcon, color = "Consumption (0.431)"))+
  geom_line(aes(y=cgdp, color = "GDP (0.453)"))+
  labs(x = "", y = "")+
  geom_text_repel(data = subset(wg_19, ccode),
                  mapping = aes(x=cum_unit, y=cgdp, label = Code), color = "red", size = 2) +
  geom_text_repel(data = subset(wg_19, ccode),
                  mapping = aes(x=cum_unit, y=cpcon, label = Code), color = "green3", size = 2) +
  scale_color_hue(name= "Weighted (2020)", direction = -1)+
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(legend.position= c(0.36, 0.84))+
  theme(legend.title = element_text(size = 8),legend.text = element_text(size = 7))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(axis.title.x =element_blank())+
  theme(text=element_text(size=8))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))

wcc21 <- ggplot(wg_21, aes(x=cum_unit)) + 
  geom_abline(color = "grey")+
  geom_line(aes(y=cpcon, color = "Consmption (0.432)"))+
  geom_line(aes(y=cgdp, color = "GDP (0.455)"))+
  labs(x = "", y = "")+
  geom_text_repel(data = subset(wg_19, ccode),
                  mapping = aes(x=cum_unit, y=cgdp, label = Code), color = "red", size = 2) +
  geom_text_repel(data = subset(wg_19, ccode),
                  mapping = aes(x=cum_unit, y=cpcon, label = Code), color = "green3", size = 2) +
  scale_color_hue(name= "Weighted (2021)", direction = -1)+
  theme(axis.text.y = element_blank(), axis.ticks = element_blank()) +
  theme(legend.title = element_text(size = 8),legend.text = element_text(size = 7))+
  theme(legend.background = element_rect(fill = "transparent"))+
  theme(axis.title.x =element_blank())+
  theme(text=element_text(size=8))+
  theme(legend.position= c(0.34, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))

wcc19_21<- ggarrange(wcc19, wcc20, wcc21, labels = c("d", "e", "f"), ncol = 3, nrow = 1,
          hjust = -1.5, font.label = list(size = 8, color = "black", face = "bold"))
```

#### put the two curves together
```{r tidy=TRUE}
cc19_21 <-ggarrange(ucc19_21, wcc19_21, ncol = 1, nrow = 2)

cc_1921 <- annotate_figure(cc19_21, left = text_grob("Cumulative % of GDP or Consumption Per Capita", rot = 90, vjust = 1, size =8), bottom = text_grob("Countries Ranked by Corresponidng Year GDP Per Capita", vjust =0, size =8))

file.plot <- file.path(dir.plots, "cc_1921.png")
png(file.plot, height=4.5, width=7, units="in", bg="white", res=300)
cc_1921
dev.off()
```

#### Kakwani Index
### Unweighted Pseudo Gini for per capita expenditure from 2000 to 2022
```{r tidy=TRUE}
library(hash)
by_year <- function(gdp_con44, yr) {
  return(gdp_con44 %>% filter(Year == yr))
}

ex_ci <- function(gdp_con44){
  ha <-hash()
  for(j in 2000:2022){
    ha[[as.character(j)]] <-by_year(gdp_con44, j)
  } 
  hb <- hash()
  keys_ha <- keys(ha)
  for(i in keys_ha) {
    bar <- calcSConc(ha[[i]][, "pcon"], ha[[i]][, "pc_gdp"], param=2)
    hb[[i]] <- bar$ineq$index
  }
  return(hb)
}
con_exp <- ex_ci(gdp_con44)
```

#### Convert hash table to dataframe
```{r tidy=TRUE}
Year <- keys(con_exp)
pgini_exp <- values(con_exp)
kd <- cbind(Year, pgini_exp)
kl <- as.data.frame(kd)
df_ci <- data.frame(lapply(kl, as.character), stringsAsFactors=FALSE)
df_ci$Year <- as.numeric(df_ci$Year)
df_ci$pgini_exp <- as.numeric(df_ci$pgini_exp)
```

#### Per capita GDP Gini (144 countries from 2015)
```{r tidy=TRUE}

dat_00_22 <- gdp_con44 %>% filter(Year>=2000)
gini_g <- dat_00_22 %>% group_by(Year) %>% summarise(gini_gdp = gini(pc_gdp), nc=n()) 
```

#### Gini of GDP and Pseudo Gini of per capita Expenditure (Kakwani Index)
```{r tidy=TRUE}
ci_ge <- left_join(df_ci, gini_g)
kak_ci <- ci_ge %>% mutate(kakwani = gini_gdp - pgini_exp)
```

#### Population Weighted Pseudo Gini for Per Capita Consumption
```{r tidy=TRUE}
ex_wci <- function(gdp_con44){
  ha <-hash()
  for(j in 2000:2022){
    ha[[as.character(j)]] <-by_year(gdp_con44, j)
  } 
  hb <- hash()
  keys_ha <- keys(ha)
  for(i in keys_ha) {
    bar <- calcSConc(ha[[i]][, "pcon"], ha[[i]][, "pc_gdp"], ha[[i]][, "pop"], param=2)
    hb[[i]] <- bar$ineq$index
  }
  return(hb)
}
wci_exp <- ex_wci(gdp_con44)
```

#### Convert hash table to dataframe
```{r tidy=TRUE}
Year <- keys(wci_exp)
wpgini_exp <- values(wci_exp)
wkd <- cbind(Year, wpgini_exp)
wkl <- as.data.frame(wkd)
df_wci <- data.frame(lapply(wkl, as.character), stringsAsFactors=FALSE)
df_wci$Year <- as.numeric(df_wci$Year)
df_wci$wpgini_exp <- as.numeric(df_wci$wpgini_exp)
```

#### Population Weighted per Capita GDP Gini
```{r tidy=TRUE}
wgini_g <- dat_00_22 %>% group_by(Year) %>% summarise(wgini_gdp = gini.wtd(pc_gdp, pop)) 
```

#### Population Weighted Gini of GDP and Pseudo Gini of per capita Expenditure (Kakwani Index)
```{r tidy=TRUE}
wci <- left_join(df_wci, wgini_g)
kak_wci <- wci %>% mutate(wkakwani = wgini_gdp - wpgini_exp)
kak_wuci <- cbind(kak_ci, kak_wci)
kakwani_wuci <- kak_wuci %>% select(1:5, 7:9)

file.output <- file.path(dir.outdata, "kakwani_wuci.csv")
write.table(kakwani_wuci, file.output, sep=",", row.names=FALSE, col.names=TRUE, quote=FALSE)
```

#### Population Weighted/unweighted kakwani Index
```{r tidy=TRUE}
WUK1 <- ggplot(kakwani_wuci, aes(x=Year)) + 
  geom_line(aes(y=wgini_gdp, color = "Per Capita GDP Gini"), linetype ="solid", size=0.8)+
  geom_line(aes(y=wpgini_exp, color = "Per Capita Consumption Pseudo Gini"), linetype ="solid", size=0.8)+
  geom_line(aes(y=wkakwani, color = "Kakwani Index"), linetype ="solid", size=0.8)+
  geom_line(aes(y=gini_gdp, color = "Per Capita GDP Gini"), linetype ="dashed")+
  geom_line(aes(y=pgini_exp, color = "Per Capita Consumption Pseudo Gini"), linetype ="dashed")+
  geom_line(aes(y=kakwani, color = "Kakwani Index"), linetype ="dashed")+
  scale_color_discrete(name = "Line Color", limits=c("Per Capita GDP Gini", 
                              "Per Capita Consumption Pseudo Gini", "Kakwani Index"))+
  labs(x = "Year",
       y = "Per Capita GDP (Consumption) Gini (Pseudo Gini)")+
  geom_line(aes(y=wgini_gdp, linetype ="solid"))+
  geom_line(aes(y=wpgini_exp, linetype ="solid"))+
  geom_line(aes(y=wkakwani, linetype ="solid"))+
  geom_line(aes(y=gini_gdp, linetype ="dashed"))+
  geom_line(aes(y=pgini_exp, linetype ="dashed"))+
  geom_line(aes(y=kakwani, linetype ="dashed"))+
  scale_linetype_discrete(name= "Line Type", limits=c( "Population Weighted","Unweighted"))+
  theme(axis.title.x = element_text(size=8)) +
  theme(axis.title.y = element_text(size=8)) +
  scale_y_continuous(limits = c(-0.05, 0.60), breaks = seq(-0.05, 0.60, 0.05)) +
  scale_x_continuous(breaks = seq(min(kak_wci$Year), max(kak_wci$Year), 1)) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  theme(text=element_text(size=8))+
  theme(legend.position= c(0.34, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))+
  theme(legend.title = element_blank())+
  theme(legend.text = element_text(size = 7))+
  theme(legend.background = element_rect(fill = "transparent")) +
  theme(legend.position= c(0.34, 0.50))
WUK1
```

#### Kakwani Index Without China, Indian and US
```{r tidy=TRUE}
nChina <- gdp_con44 %>% filter(Country !="China")
nIndia <- gdp_con44 %>% filter(Country != "India")
nUS <- gdp_con44 %>% filter(Country !="United States")
```

#### Gini without China, India or US
```{r tidy=TRUE}
gini_nchina <- nChina %>% group_by(Year) %>%
  summarise(gini_nC = gini.wtd(pc_gdp, pop))

gini_nindia <- nIndia %>% group_by(Year) %>%
  summarise(gini_nI = gini.wtd(pc_gdp, pop))

gini_nUS <- nUS %>% group_by(Year) %>%
  summarize(gini_nUS = gini.wtd(pc_gdp, pop))

all_gini <- gdp_con44 %>% group_by(Year) %>%
  summarise(g_all = gini.wtd(pc_gdp, pop))

wtgini_all <- list(all_gini, gini_nchina, gini_nindia, gini_nUS) %>% 
  reduce(left_join, by = "Year")
```

#### Consumption concentration without China, India or US 
```{r tidy=TRUE}
ci_nchina <- nChina %>% group_by(Year) %>%
  summarise(ci_nC = calcSConc(pcon, pc_gdp, w = pop, param = 2)$ineq$index)

ci_nindia <- nIndia %>% group_by(Year) %>%
  summarise(ci_nI = calcSConc(pcon, pc_gdp, w = pop, param = 2)$ineq$index)

ci_nUS <- nUS %>% group_by(Year) %>%
  summarise(ci_nUS = calcSConc(pcon, pc_gdp, w = pop, param = 2)$ineq$index)

all_ci <- gdp_con44 %>% group_by(Year) %>%
  summarise(ci_all = calcSConc(pcon, pc_gdp, w = pop, param = 2)$ineq$index)

wtci_all <- list(all_ci, ci_nchina, ci_nindia, ci_nUS) %>% 
  reduce(left_join, by = "Year")

kakci_gini <- cbind(wtgini_all, wtci_all)
kakci_gini <- kakci_gini %>% select(1:5, 7:10)
```

#### Kakwani Index Calculation
```{r tidy=TRUE}
kci <- kakci_gini %>% mutate(k_all = g_all- ci_all, k_nC = gini_nC - ci_nC, 
                             k_nI = gini_nI - ci_nI, k_nUS = gini_nUS - ci_nUS)
kci <- kci %>% select(1, 10:13) 

file.output <- file.path(dir.outdata, "kci.csv")
write.table(kci, file.output, sep=",", row.names=FALSE, col.names=TRUE, quote=FALSE)
```

#### Kakwani Index Calculation
```{r fig.height = 3.5, fig.width = 7, fig.align = "center"}
kak_index <- ggplot(kci, aes(x = Year))+
  geom_line(aes(y = k_all, color = "All"), linetype = "solid", size = 0.8)+
  geom_line(aes(y = k_nC, color = "Without China"), linetype = "solid")+
  geom_line(aes(y = k_nI, color = "Without India"), linetype ="solid")+
  geom_line(aes(y = k_nUS, color = "Without US"), linetype = "solid")+
  scale_colour_manual(values=c("Without China" = "blue", "Without India" = "red", 
                               "Without US" = "green", "All" = "black"))+
  guides(color=guide_legend(nrow=2))+
  labs(x = "Year", y = "Weighted Kakwani Index")+
  theme(axis.title.x = element_text(size=8)) +
  theme(axis.title.y = element_text(size=8)) +
  scale_x_continuous(breaks = seq(min(kci$Year), max(kci$Year), 1)) +
  scale_y_continuous(limits = c(-0.03, 0.07), breaks = seq(-0.03, 0.07, 0.01)) +
  theme(axis.text.x = element_text(angle = 90, vjust=0.5))+
  theme(legend.title = element_blank())+
  theme(legend.direction = "horizontal") +
  theme(legend.position = c(0.3,0.88)) +
  theme(legend.box = "vertical")+
  theme(text=element_text(size=8))+
  theme(legend.position= c(0.34, 0.84))+
  theme(legend.key.width = unit(0.5, 'cm')) +
  theme(legend.key.size= unit(0.5, "cm")) +
  theme(legend.key.height = unit(0.4, "cm"))+
  theme(legend.spacing.y = unit(0, "lines"))+
  theme(legend.spacing.x = unit(0.5, "lines"))+
  theme(legend.background = element_rect(fill = "transparent"))

kak_wo <-ggarrange(WUK1, kak_index, labels = c("a", "b"), hjust = - 1.5, 
                     font.label = list(size = 8, color = "black", face = "bold"), 
                     ncol = 2, nrow = 1)

file.plot <- file.path(dir.plots, "kak_wo.png")
png(file.plot, height=3.5, width=7, units="in", bg="white", res=300)
kak_wo
dev.off()
```
